\section{Transformation of Pointing Offsets and Corrections}

Table \ref{tab:offset_and_corrections} show examples of pointing offsets and corrections applied during the pointing scans.
The column "Original" show raw pointing scan data.
The pointing corrections $ca$ and $ie$ are normally updated according to equations \eqref{eq:ca} and \eqref{eq:ie}, 
as shown in the first two rows of the table.
However, observers may choose not to update the pointing, particularly when the pointing offset is small, as illustrated in the consecutive row of the table.
In other cases, the corrections may be updated but not according to equations \eqref{eq:ca} and \eqref{eq:ie}.
This can occur when a new science project is loaded and the pointing correction from the previous time that project was used is applied.
These factors introduce several challenges for the training of a model:
\begin{itemize}
    \item $ca$ and $ie$ should represent the optimal correction using all the information we have about the current state of a system.
    If we do not update the corrections, there is some information about the system (the previously observed pointing offset) that the model is not receiving.
    \item Some features are constructed as the change in variables since the last correction.
    If the corrections are not updated, this interval is longer than if they were, and the resulting features could be more prone to uncertainties and noise.
    A problem with the integration also occurs if the corrections are not updated according to the equations \eqref{eq:ca} and \eqref{eq:ie}.
    Then, we do not know when those corrections represent the system, resulting in inaccurate features.
\end{itemize}

A possible solution to this problem is a two step procedure where we first transform the offsets and corrections to represent the system at the most recent pointing scan,
and second use these transformed offsets as training labels and the transformed corrections as training inputs.

In practice, we do this by assuming the corrections $ca$ and $ie$ are updated after every pointing scan.
This changes the correction applied during the next pointing scan, which further affects the observed pointing offset.
This effect propagates throughout the whole dataset.
The following formulas 
\begin{align}
    \Tilde{ca}_i &= \Tilde{ca}_{i-1} + \Tilde{\delta}_{az,i-1}\label{eq:ca_tilde}\\
    \Tilde{ie}_i &= \Tilde{ie}_{i-1} - \Tilde{\delta}_{el,i-1}\label{eq:ie_tilde}\\
    \Tilde{\delta}_{az,i} &= \delta_{az,i} + ca_i - \Tilde{ca}_i\label{eq:off_az_tilde}\\
    \Tilde{\delta}_{el,i} &= \delta_{el,i} - ie_i + \Tilde{ie}_i\label{eq:off_el_tilde}
\end{align}
Where the "$\sim$" denotes a transformed variable.
Using these transformations,
the corrections used for training and the resulting offset are similar to the ones that would be observed if the corrections were made according to the equations
\eqref{eq:ca} and \eqref{eq:ie} after every pointing scan. The column "Transformed" in the table shows the transformed variables.

\begin{table}[H]
    \centering
    \caption[Original and transformed pointing offsets and corrections]{\textbf{Original:} Example from the dataset of the observed pointing offsets and the corrections applied during the pointing scan.
    \textbf{Transformed:} Pointing offsets and corrections according to equations \eqref{eq:ca_tilde}, \eqref{eq:ie_tilde}, \eqref{eq:off_az_tilde}, and \eqref{eq:off_el_tilde}.}
    \label{tab:offset_and_corrections}
    \begin{tabular}{c cccc c cccc}
    \toprule
    \multicolumn{1}{c}{} & \multicolumn{4}{c}{Original} & \multicolumn{1}{c}{} & \multicolumn{4}{c}{Transformed} \\
    \cmidrule(lr){2-5} \cmidrule(lr){7-10} 
    $i$ &  $\delta_{az}$ &  $\delta_{el}$ &  $ca$ & $ie$  & & $\Tilde{\delta}_{az}$ &  $\Tilde{\delta}_{el}$ &  $\Tilde{ca}$ &  $\Tilde{ie}$ \\
    \midrule
    $1$ &     $1.2$ & $0.1$ & $2.1$ &  $1.7$ &   &    $1.2$ &       $0.1$ &       $2.1$ &       $1.7$ \\
    $2$ &     $0.0$ & $0.5$ & $3.3$ &  $1.6$ &   &    $0.0$ &       $0.5$ &       $3.3$ &       $1.6$ \\
    $3$ &    $-1.1$ & $0.0$ & $3.3$ &  $1.6$ &   &   $-1.1$ &      $-0.5$ &       $3.3$ &       $1.1$ \\
    $4$ &     $0.6$ & $0.7$ & $2.2$ &  $1.6$ &   &    $0.7$ &       $0.7$ &       $2.2$ &       $1.6$ \\
    $5$ &     $0.9$ & $1.4$ & $2.2$ &  $1.6$ &   &    $0.2$ &       $0.7$ &       $2.8$ &       $0.9$ \\
    $6$ &     $1.0$ & $1.1$ & $2.2$ &  $1.6$ &   &    $0.1$ &      $-0.3$ &       $3.1$ &       $0.2$ \\
    $7$ &    $-0.9$ & $1.2$ & $3.1$ &  $0.5$ &   &   $-1.0$ &       $1.3$ &       $3.2$ &       $0.5$ \\
    $8$ &     $0.5$ & $1.5$ & $2.2$ & $-0.7$ &   &    $0.5$ &       $1.4$ &       $2.2$ &      $-0.7$ \\
    $9$ &    $-0.3$ & $0.4$ & $2.2$ & $-0.7$ &   &   $-0.8$ &      $-1.1$ &       $2.7$ &      $-2.2$ \\
    \bottomrule
    \end{tabular}
\end{table}

\section{Feature Engineering for Transformed Offsets and Corrections}
There are two main features engineered for this project; features that represent the system during a pointing scan and features that represent changes since the last correction. 
The idea behind this is simple. The correction used during a pointing scan represents the ideal correction for the system during the previous pointing scan.
As there are a lot of factors and complex relationships, and we do not have large amounts of training data, it might be easier for the model to learn 
how these changes affect the pointing rather than learning all the relationships.

\subsection{Feature Transformation} 
\subsubsection{Median values}
The median value of variables during a pointing scan is the most used feature.

\subsubsection{Sum of all change}
To capture systematic error in pointing due to the telescope moving back and forth in azimuth and elevation,
we sum over the positive and negative changes in these variables.

Given the time of the last pointing correction $t_1$ and the start of a pointing scan $t_2$, the sum over the positive changes in a variable $x_i$ is given by
\begin{equation}\label{eq:positive_int}
    X = \sum_{i=t_1+1}^{t_2} \max(0, x_i-x_{i-1})
\end{equation}

Similarly, the sum of negative changes in a variable is
\begin{equation}\label{eq:negative_int}
    X = \sum_{i=t_1+1}^{t_2} \min(0, x_i-x_{i-1})
\end{equation}

We make these features with azimuth and elevation.

\subsubsection{Change since the last correction}
This feature is self-explanatory and is just the change in a variable since the pointing was corrected.
\begin{equation}
    \Delta x = x_{t_2} - x_{t_1}
\end{equation}

In order to make this feature more robust against noisy data,
we instead consider the change in the median for a time interval around the last correction $t_1$ and the start of a pointing scan $t_2$

\begin{equation}
    \Delta x = \textit{median}(x_{t_2}, x_{t_2 - 1}, \dots, x_{t_2- p}) - \textit{median}(x_{t_1}, x_{t_1 + 1}, \dots, x_{t_1 + p}),
\end{equation}
where $p$ is the number of data points needed to cover a period of $P$ minutes, given by $p = P \cdot frequency$. The unit of frequency is data points per minute, found in Table \ref{tab:data_frequency}.

\subsubsection{Max change in time interval}
In case the speed of the temperature change affects the deformation of the telescope's structure, we find the maximum temperature change in a given time interval since the last pointing correction.

\begin{equation}
    X = \max (x_{t_1+p} - x_{t_1}, x_{t_1+p} - x_{t_1}, \dots, x_{t_2} - x_{t_2-p}),
\end{equation}


\subsubsection{Position of the sun}
Observers at the telescope report that the sun is affecting the pointing.
It is most drastically affected when the sun sets or rises, likely due to rapid temperature change leading to deformation in the telescope structure.
We also think the sun's position affects the pointing.
For instance, if the sun is shining on the left side of the telescope, it will affect the pointing differently than if it is on the right side.
Obtaining the sun's position for the telescope's location is done using the python module PyEphem \cite{ephem}.

Using the azimuth angle of the sun and the telescope, we can calculate the position of the sun with respect to the pointing with
\begin{equation}
    \Delta \textit{Az}_\odot = \textit{Az}_{\textit{t}} - \textit{Az}_\odot
\end{equation}
This will result in values outside the $[-180^\circ,180^\circ]$. An example is if $Az_\odot=179^\circ$ and $Az_t = -179^\circ$.
The calculation in equation \eqref{eq:sun_az_diff} yield $-179^\circ-179^\circ=-358^\circ$,
which corresponds to the sun being $358^\circ$ to the right of the telescope, while it ideally should be $2^\circ$ to the left.
Therefore, we adjust the values accordingly
%Write equations where \Delta Az_\odot above 180 is -= 360 and under 180 is += 360
\begin{align}
    \Delta Az_\odot = Az_\odot +360^\circ, \; \text{for} \; \Delta Az_\odot < 180^\circ\\
    \Delta Az_\odot = Az_\odot -360^\circ, \; \text{for} \; \Delta Az_\odot > 180^\circ
\end{align}

Here, the interval of the difference in azimuth is fixed to the interval $(-180^\circ,180^\circ)$,
where $0^\circ$ means the telescope is pointing towards the sun in the azimuth direction.
$\Delta \textit{Az}_\odot = 90^\circ$ corresponds to the sun being direct to the left of the pointing direction. \\

Another measure tested is the total angle between the pointing and the sun's position. We calculate this using the following formula
\begin{equation}
    \theta = \cos \textit{Az}_t \cdot \cos \textit{El}_t\cdot \cos \textit{Az}_\odot \cdot \cos \textit{El}_\odot + \sin \textit{Az}_t \cdot \cos \textit{El}_t\cdot \sin \textit{Az}_\odot \cdot \cos \textit{El}_\odot + \sin \textit{El}_t \cdot \sin \textit{El}_\odot
\end{equation}