\section{Experiment 1: Pointing Correction Model}
The first research question addressed in this thesis is whether machine learning can enhance the pointing accuracy of a radio telescope using the current pointing strategy.
To investigate this question, we explore a realistic scenario (case 1) in which a model is trained on a smaller period of data and used to predict the offset of consecutive scans for a period afterward.
We focus now on the model predicting the offsets of only NFLASH230. 
The results from this case, presented in Table \ref{tab:results_minval_val_test_days_04_n230}, demonstrate that the model's performance on the validation set is promising,
with the root-mean-square (RMS) ratio in the range of approximately $0.80$-$0.85$ for azimuth and elevation,
which corresponds to a $15$-$20\%$ reduction in pointing offset.
However, this performance does not transfer to the following test period,
in which the RMS ratios are $1.16$-$1.46$, indicating a $16$-$46\%$ increase in pointing offset.
There are several possible reasons for the model's poor performance on the test set.
One of the limitations of tree-based models, such as XGBoost, is that they typically do not generalize well to new data that is different from the training data,
as they predict solely on logical conditions seen in the training set.
If the factors affecting the pointing offset change over time and the new data is very different from the training data, the model will likely perform poorly. 
Furthermore, another potential explanation for the poor performance could be that the dataset is too small, and the model overfits on the validation set.
The results of this experiment suggest that learning the relationships in the data that affect pointing offset is challenging, and a complex model may be necessary.
To train a proper complex model, a larger amount of data is required, at least more than the number of samples in the training and validation sets for case 1, being \textcolor{red}{add samples here}.
The findings also indicate that choosing the complexity of the model with the best performance on the validation set may not necessarily lead to the best performance on the test set.
We further explore this aspect by examining Table \ref{tab:results_nflash_days}, which demonstrates the mean RMS ratio on the test set using the same number of features for all the folds.
This provides an idea of the complexity that might provide the best performing models on the test set.
Even though the model with the best performance on the validation set is not chosen, which could potentially be a lucky performance or overfitted,
no improvement is observed in the current pointing model on the test set.
However, the results show better performance than the first Table \ref{tab:results_minval_val_test_days_04_n230}.
The same trends are observed when predicting offsets from all instruments in Tables \ref{tab:results_minval_val_test_days_04_all} and \ref{tab:results_all_days}. \\

Moving on to case 2, which tests whether the amount of data is a limitation for enhancing the pointing accuracy.
This test case is less realistic because the data is split into $6$ folds and cross-validation is performed,
and for all folds except for the last one, the test period will be either before or between the training/validation set, making the results less applicable in practice.
Results from this case indicate that a larger time period helps the model generalize better, which is expected as a larger period includes more variation that can help the model capture the relationships between features.
We start by looking at the same table \ref{tab:results_minval_val_test_days_04_n230}.
Here, we also see a good performnce on the on the validation set across all folds, with a $9$-$20\%$ reduced pointing offset on the validation set.
This shows that more training data helps the model generalize, which is somewhat expected considering that a larger time period includes more variation,
which helps the model learn what causes the offsets.
The average RMS ratio over all folds on the test set for case 2 is $0.958$ for azimuth and $0.941$ for elevation.
With the standard deviations, the $95\%$ confidence intervals would be put at $[0.850, 1.066]$ for azimuth and $[0.786, 1.010]$ for elevation.
Given that the upper bound of both confidence intervals are larger than $1$, and the fact that the testing case is not realistic, we cannot conclude that the model can reduce the pointing offset in a robust and consistent manner.
For an unbiased result that reflects expected performance in practice, the RMS ratio for the last fold in Table \ref{tab:results_minval_val_test_days_04_n230} is used,
showing a $7.0\%$ and $9.4\%$ reduced RMS for azimuth and elevation, respectively.
This indicates that a possible pointing strategy could be training a model on multiple months worth of data and then using the model for a couple of weeks.
However, given the limited data available in this project, this could not be tested thoroughly.
If more data were available, a similar analysis could be performed with the start and end time of both the train/validation and test set moved by two weeks,
and iteratively train new models to predict the offsets for the next time period.
This could verify whether the improved performance repeats.\\

In case 2, the model predicting offsets from all instruments shows results similar to the model predicting only NFLASH230 offsets, though with slightly worse performance.
The reduced performance is likely due to less of training data available for the other instruments.
The difference in performance between these two models provides some insights.
First, it suggests that various factors influence the different instruments differently.
Otherwise, we expect the larger model to perform better than the smaller.
The reason that each instruments are affected differently can be the different mounting locations at the telescope.
Different mounting locations mean that the path of photons leading to the receiver can vary, leading to distinct factors affecting the pointing.\\


There are several measures we could take to enhance the analysis further. One such measure is to improve feature engineering.
When the quantity of training data is limited, it is vital to ensure that the features are as informative as possible.
For example, instead of including six distinct temperature measurements in a model,
it may be preferable to create terms based on the difference between multiple temperature measurements, as they did in \cite{whitegreen2022}.

Another step is to integrate corrections for $ca$ and $ie$ into the offset that we intend to predict.
These corrections are applied at each pointing scan and display the strongest correlation to the target offset of all the features,
and we therefore incorporated them in all the models.
By transforming the offsets $\delta_{az}=\delta_{az} + ca$ and $\delta_{el}=\delta_{el} + ie$, we can then eliminate $ca$ and $ie$ from the input to the models.
This could remove a layer of complexity for the model.\\


Another option to consider for predicting the pointing offsets is to use neural networks.
These models offer several advantages if trained successfully.
For instance, they can handle multiple outputs and thus only require a single model.
By training the neural network with two outputs, it can simultaneously consider the offsets in azimuth and elevation, and explore the correlation between them.
It would also be beneficial to fine-tune the network as new pointing scans become available continuously.
Despite their potential benefits, our initial tests with neural networks did not yield satisfactory results,
likely due to the limited amount of training data available.\\


To further improve the analysis, there are other possible areas for exploration, such as minimizing the number of pointing scans conducted by the astronomers.
While performing scans every one to two hours is standard practice, this can be time-consuming and may disrupt science observations.
A possible approach to investigate this possibility is to identify periods when pointing scans are conducted every hour or so.
Rather than providing the model with the corrections $ca$ and $ie$ obtained following the previous pointing scan, the model can be fed with the corrections from a pointing scan that occurred six hours earlier to evaluate its performance.
In this case, reducing pointing offsets is not a primary objective of the machine learning approach.
Instead, maintaining the same level of pointing accuracy while conducting fewer pointing scans would be a satisfactory outcome.\\



% \begin{itemize}
%     \item comment on parameters
%     \item discuss the adjustments made by system,a nd that the analytical poinitng models change over time too, which can make it harder for amodel, considering there are many other factors to ocnsider.
% \end{itemize}

% \section{Experiment 2: Pointing Correction Model}
% The first research question is if we can use machine learning to increase the pointing accuracy using the same pointing strategy as today.
% We start by discussing case 1, which test a realistic scenario of training a model on a smaller period,
% and then using the model to predict the offset of the consecutive scans for a period.
% We look first at the model using NFLASH230 data only.
% The results from this case show that the performance on the validation set is good \ref{tab:results_minval_val_test_days_04_n230},
% with the RMS ratio being in the range $\tilde{}0.80$-$0.85$ for azimuth and elevation.
% That corresponds to a $15$-$20\%$ reduced pointing offset.
% This performance does not transfer to the following test period,
% for which the RMS ratios are in the range $1.16$-$1.46$, corresponding to a $16$-$46\%$ increased pointing offset.
% There could be multiple reasons for the drastically worse performance set.
% One possibility is that the relationships learned by the model change over time, therefore the model is not able to perform well.
% This is one limitation of tree based models like XGBoost.
% Because the model is predicting solely on logical conditions seen in the training set,
% it is not able to generalize well to new data.
% A neural network could potentially perform better on unseen data, as it train a continuous function.
% Another reason is that the dataset is too small, and the performance on the validation set is based on luck.
% In other words, model overfits on the validation set.
% From this project, it is clear that learning the relationships in the data that affects the pointing offset is not easy, and a complex model might be essential.
% And in order to train a proper complex model, we need large amounts of data,
% at least more than the \textcolor{red}{train set num} and \textcolor{red}{train set num} samples in the training and validation set respectively.
% The resutls we just discussed, uses a strategy where we choose the complexity of the model with the best performance on the validation set.
% We now look at Table \ref{tab:results_nflash_days}, which shows the mean RMS ratio on the test set using the same number of features $k$ for all the folds.
% This gives an idea on what complexity could provide the best models.
% It shows that even though we do not choose the model with the best performance on the validation set, which could potentionally be a lucky performance or overfitted,
% we still do not see any improvement of the current pointing model on the test set.
% We do however see better results than in the first table \ref{tab:results_minval_val_test_days_04_n230}.
% If we look at the model using the data from all instruments, we see similar results \ref{tab:results_minval_val_test_days_04_all} \ref{tab:results_all_days}.\\

% Now onto case 2, which show a less realistic test case, but can give an idea on if the amount of data is a limitation.
% This test case is less realistic because we split the data into $6$ folds and perform cross validation.
% For all folds except for the last one, the test period will be either before, or between the training/validaiton set.
% Therefore, these results does not reflect expected performance in practise.
% We start by looking at the same table \ref{tab:results_minval_val_test_days_04_n230},
% with results from the NFLASH230 model choosing the best model from the validation set, and tests the performance on the test set.
% Again, we see a good performance on the validation set across all folds, with a $9$-$20\%$ reduced pointing offset, for azimuth and elevation models.
% Now, we also see a reduced pointing offset even on the test set.
% This shows that more training data helps the model generalize,
% which is somewhat expected considering that a larger time period includes more variation, which helps the model capture relations between features.
% The average RMS ratio over all folds for case 2 is $0.958$ for azimuth and $0.941$ for elevation.
% With the standard deviations, the $95\%$ confidence intervals would be put at $[0.850, 1.066]$ for azimuth and $[0.786, 1.010]$ for elevation.
% The results show that there is something to be learnt in the data, and this pointing strategy could potentially lead to improved pointing.
% For a completely unbiased result, which show a performance that could have been achieved in practice,
% we look at the RMS ratio for the last fold in Table \ref{tab:results_minval_val_test_days_04_n230}.
% This shows a $7.0\%$ and $9.4\%$ reduced RMS for azimuth and elevation respectively,
% which indicates that a possible pointing strategy could be training a model on multiple months worth of data, and then use the model for a couple of weeks.
% Given we only have six months of data, this could not be tested more thoroughly.
% If we had more data, we could trained a model on the same amount of data, but moving the start and end time of both the train/validation and test set by two weeks, and
% see if the improved performance repeats.




% \begin{align}
%     CI_{Az} &= \bar{r}_{RMS_Az} \pm 1.96 \cdot \sigma_Az = [0.958 - 1.96\cdot 0.055, 0.958 + 1.96\cdot 0.055] = [0.850, 1.066]  \\
%     CI_{El} &= \bar{r}_{RMS_El} \pm 1.96 \cdot \sigma_El = [0.941 - 1.96\cdot 0.079, 0.941 + 1.96\cdot 0.079] = [0.786, 1.010]
% \end{align}


\section{Experiment 2: Pointing Model using Neural Networks}
To address the second research question of whether machine learning models can replace the traditional analytical linear regression models commonly used in radio/(sub-)mm telescopes,
we utilized raw data from the apex telescope containing input and observed coordinates for azimuth and elevation.
Typically, linear regression models are fitted to this raw data using multiple geometrical and empirical terms.
As a benchmark, we used a linear regression model and compared its performance against four different neural network architectures (see Figure \ref{fig:nn_architecture}).
Our goal was to determine if a machine learning approach could outperform the linear regression model and, if so, which type of architecture performs the best.
We employed cross-validation with six folds and found that all neural network architectures significantly outperformed the linear regression model.
The mean RMS of the linear regression model was $47.63''$ with a standard deviation of $12.81''$, whereas all neural network architectures fell within the range of $16''$-$19''$ with deviations of $5''$-$7''$.
Although all the neural network architectures had similar performance and standard deviation, we cannot conclude that any of them is better than the others.
Overall, our results indicate that machine learning models can outperform traditional analytical models for pointing offset prediction in radio telescopes,
and we encourage further research in this area to explore the full potential of machine learning techniques.\\

Analysis of the hyperparameters selected for the top-performing models reveals that most of them are relatively simple in structure.
Interestingly, none of the models employed two hidden layers.
Additionally, the architectures incorporating distinct layers for the terms employed in the analytical pointing models exhibited lower neuron counts in their nonlinear layers.
This is likely because the level of complexity required for the input features provided to the nonlinear layers is assumed to be lower when other aspects of the model are responsible for handling certain factors.
Thus, the observed pattern of hyperparameter selection is consistent with the expectation based on architecture.\\

Table \ref{tab:exp1_features} presents the features selected by each model.
All the models included COMMANDAZ and COMMANDEL as inputs for the nonlinear layers. The "Separated" neural networks incorporated the harmonic features as inputs in the linear part of the network.
We sampled the remaining features randomly.
Interestingly, the models used few nonlinear features, indicating that the factors influencing pointing error are difficult to model with the current training set.
Obtaining more data may be necessary to establish more complex relationships between features that can improve the performance of the pointing models.
Notably, the regular neural network utilized only one additional feature, namely DISP\_ABS (which still requires explanation).
This finding suggests that a neural network utilizing only the azimuth and elevation angle could outperform a linear regression model.
However, linear regression models have advantages in terms of transparency and robustness, as they are inherently linear, and all the features in the model are independent of each other,
resulting in a model that is unlikely to result in significant pointing errors.
This is an important consideration when using machine learning models, which can be challenging to interpret.
Their robustness must be thoroughly tested before being deployed in practical settings.
These considerations also underlie the design of the "Separated" neural network architectures.
We wanted to utilize the robustness of linear regression on the terms used in the current analytical pointing model while incorporating new features. 
Considering that the experiment demonstrated negligible performance differences between the separate neural networks and the regular
one and that the best models only choose a few features, we conclude that additional training data is required to see if this is a feasible approach.\\

While the experiment did not demonstrate the performance that would be expected from using such models in practice,
it indicates that a neural network could provide a more efficient and effective model than a linear regression model.
Moreover, a neural network can be fine-tuned with newly obtained data, making it an appealing model.
If this approach is shown to be reliable and high-performing, it could be the preferred choice for more advanced telescopes, as it may require less manual analysis to develop and maintain a pointing model.
However, further testing with additional data is needed to confirm the feasibility of this approach.
